Analytical Predictions of Explanet Obliquities Generated by Planet-Disk Interactions

### 2019-10-24 writeups/Su_disk_disp.tex
okay full draft kinda

    \documentclass[
        fleqn,
        usenatbib,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        ae, aecompl,
        graphicx,
        booktabs,
    }
    \usepackage[T1]{fontenc}
    \usepackage[
        labelfont=bf, % caption names are labeled in bold
        font=scriptsize % smaller font for captions
    ]{caption}
    \usepackage[font=scriptsize]{subcaption} % subfigures
    \captionsetup{compatibility=false}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

\title[Cassini State Capture]{Cassini State Capture}
\author[Y. Su et\ al.]{
Yubo Su$^1$,
Dong Lai$^1$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
}

@@

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\renewcommand*{\sectionautorefname}{Section}
\maketitle

\begin{abstract}
    Abstract
\end{abstract}

@@

The primary result of our paper is \autoref{fig:nonad_3_ensemeble}.

@@

Denote $\hat{s}$ spin of planet, $\hat{l}$ angular momentum of planet, and
$\hat{l}_d$ angular momentum of the surrounding disc. We approximate $S \ll L
\ll L_d$, so $\hat{l}_d$ is approximately constant and $\hat{l}$ experiences
negligible backreaction torque from $\hat{s}$. Much of this treatment borrows
from \citealp{anderson2018teeter} and \citealp{millholland_disk}. We consider
precession equations
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}}\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, $M_\star$ is
the mass of the host star, and $M_p, R_p, a_p, \Omega_p$ are the mass, radius,
semi-major axis and angular spin frequency of the planet respectively.
\autoref{eq:wld} is written for a disk of mass $M_d$ located at characteristic
distance $r_d$ from the host star \citep[see][for a power-law disk
profile]{millholland_disk}, $n \equiv \sqrt{GM_\star/a_p^3}$ is the planet's
orbital mean motion. The three angles formed by the three vectors are notated
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

Let us transform to frame rotating about $\hat{l}_p$ with frequency $g \equiv
-\omega_{ld}\cos I < 0$, where $\hat{l}$ fixed. In this frame, $\hat{l}_p$
also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_d},
\end{equation}
where $\alpha \equiv \omega_{sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta \equiv{}& \frac{\abs{g}}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p}{0.001 M_\star}}
            \p{\frac{a_p}{1\;\mathrm{AU}}\frac{R_J}{R_p}}^3
            \p{\frac{2\pi/\p{10\;\mathrm{days}}}{\Omega_p}}\nonumber\\
        &\p{\frac{M_d}{0.01 M_{\star}}}
            \p{\frac{30 a_p}{r_d}}^3
            \p{\frac{n}{2\pi / \p{1\;\mathrm{yr}}}}
            \p{\frac{\cos I}{\cos 5^\circ}},\nonumber\\
    \tau \equiv{}& \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ as well. We nondimensionalize:
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_d},\\
        ={}& 0.015 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p}{0.001 M_\star}}
            \p{\frac{a_p}{1\;\mathrm{AU}}\frac{R_J}{R_p}}^3\nonumber\\
        &\p{\frac{2\pi/\p{10\;\mathrm{days}}}{\Omega_p}}
            \p{\frac{10\;\mathrm{Myr}}{t_d}},\\
    \rd{\eta}{\tau} ={}& -\epsilon \eta.\label{eq:deta_dt}
\end{align}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

For analytical work, we adopt spherical coordinate system where $\hat{l} =
\hat{z}$ and $\theta, \phi$ are the polar and azimuthal angle of $\hat{s}$. By
convention, we choose $\hat{l}_z$ at coordinates $\theta = I, \phi = \pi$.
Note that $\p{\cos \theta, \phi}$ form a canonically conjugate pair of variables
describing $\hat{s}$. To derive equations of motion for these canonical
coordinates, we introduce Hamiltonian
\begin{align}
    \mathcal{H}\p{\theta, \phi} &= -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_d}.\label{eq:H}\\
        &= -\frac{1}{2}\cos^2\theta
            + \eta \p{\cos \theta \cos I - \sin I \sin \theta \cos \phi}.
\end{align}

The equations of motion follow by taking derivatives of $\mathcal{H}$ and agree
with \autoref{eq:dsdt_base}:
\begin{subequations}\label{se:H_eom}
    \begin{align}
        \pd{\phi}{t} = \pd{\mathcal{H}}{(\cos\theta)}
            &= -\cos\theta + \eta\p{\cos I + \sin I \cot \theta \cos \phi},\\
        \pd{(\cos \theta)}{t} = -\pd{\mathcal{H}}{\phi}
            &= -\eta \sin I \sin \theta \sin \phi.
    \end{align}
\end{subequations}
While these equations are numerically stiff owing to the $\cot\theta$ term, they
are the most intuitive description for analytical work.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(this follows from linearization of \autoref{se:H_eom}). When $\eta < \eta_c$,
where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$. Note that
    $\theta \in [-\pi, \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown in \autoref{fig:eq_1contours}.
Note that trajectories in zone $II$ librate about CS2 while those in zones
$I$/$III$ circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$, where warmer
    colors denote more positive values. The black solid line is the separatrix,
    which only exists for $\eta < \eta_c$. The three zones, divided by the
    separatrix, are labeled. The Cassini states are labeled and have the same
    colors as \autoref{fig:cs_locs}. The interior of the separatrix, shaded in
    grey, is formally only defined for $\eta < \eta_c$, but we may identify the
    phase space trajectories that flow into zone $II$ when evolved forward in
    time; this is the shaded region in the top left plot, bounded by the dotted
    black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$. This extension is
also reflected in the dashed boundary in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional $A_{i}(\eta)$ as given by
    \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{s:app_transition}. Dashed lines for
    $\eta > \eta_c$ effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$,
    denoting the area of points that would flow into either area under adiabatic
    decrease of $\eta$ to below $\eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004I}. The resonant
libration period is just the libration period about CS2, which can be found by
linearizing \autoref{se:H_eom} about CS2. This results in constraint
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{1}{T_{lib}},\nonumber\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin\theta \sin I
            \p{1 + \eta_\star \frac{\sin I}{\sin^3\theta}}}.\label{eq:ad_constr}
\end{align}
This formula differs from that presented in \citealt{millholland_disk}.

@@

We evenly sample $101$ values of $\theta_{sd, i}$, and for each $\theta_{sd, i}$
value, we pick $101$ $\hat{s}$ initial conditions approximately on the ring
fixed by $\theta_{sd, i}$\footnote{For each $\theta_{sd, i}$ we choose an
initial condition on the libration cycle with initial condition $\p{\theta_2 +
\theta_{sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates of
CS2), such that all initial conditions for a given $\theta_{sd, i}$ enclose the
same initial phase space; see \autoref{sss:a_evo}.}. To be concrete, we choose
$\eta_i = 10\eta_c$. For each initial condition, we evolve
Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt} until $\eta = 10^{-5}$. For
such small $\eta$, $\hat{s}$ precesses about $\hat{l}$ with uniform final
obliquity $\theta_f$. This transfer function $\theta_f\p{\theta_{sd, i}}$ is our
primary result, and is shown for $I = 5^\circ, 10^\circ, 20^\circ$ in
\autoref{fig:ad_ensemble}, \autoref{fig:3_ensemble_10_35}, and
\autoref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the simulation, which are in excellent agreement with the colored tracks.
These colored tracks are calculated semi-analytically using methods explained in
the following section.

>>>>

    \documentclass[
        fleqn,
        usenatbib,
        referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

\title[Analytical Exoplanet Obliquities]{Analytical Predictions of Explanet
Obliquities Generated by Planet-Disk Interactions}
\author[Y. Su et\ al.]{
Yubo Su$^1$,
Dong Lai$^1$
\\
$^1$ Cornell Center for Astrophysics and Planetary Science, Department of
Astronomy, Cornell University, Ithaca, NY 14853, USA
}

@@

\begin{document}\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\renewcommand*{\sectionautorefname}{Section}
% \renewcommand*{\subsectionautorefname}{Subsection}
\maketitle

\begin{abstract}
    Large planetary spin-orbit misalignments (obliquities) are thought to play a
    key role in some planets' atmospheric circulation and tidal heating. A
    promising avenue to generate high obliquities is via spin-orbit resonances,
    where the spin and orbital precession frequencies of the planet become and
    remain commensurate as the system evolves. One such proposed mechanism
    involves a dissipating protoplanetary disk driving orbital precession of an
    interior planet \citep{millholland_disk}. This previous work is limited to
    small initial misalignment angles between the spin axis of the planet and
    orbital axis of the massive disk. In the present paper, this scenario is
    analytically characterized and extended to arbitrary misalignment angles. It
    is shown that (i) under adiabatic evolution, final planetary obliquities
    bifurcate into distinct tracks governed by interactions with the resonance,
    while (ii) under non-adiabatic evolution, a broad range of obliquities can
    be excited. Analytical formulae are provided for most interesting regimes of
    parameter space.
\end{abstract}

@@

Substantial planetary obliquities, defined as the angle between a planet's spin
and orbital axes, have featured in many astrophysical scenarios. They lie at the
heart of the obliquity tides scenario of radius enhancement in hot Jupiters
\citep{winn_otides, fabrycky_otides, millholland_wasp12b}. They have also been
invoked to explain peculiar thermal phase curves in hot Jupiters \citep[see
e.g.][]{millholland_signatures, ohno_infer_obl}. While exoplanet obliquities
have yet to be directly detected, techniques have been proposed to constrain
them \citep[see e.g.][]{schwarz_infer_obl, rauscher_infer_obl}.

Many mechanisms have been proposed to generate large planetary obliquities.
Giant impacts are traditionally used to generate the large obliquities of the
solar system ice giants \citep{original_gi, morbidelli_gi}. More recently,
obliquity excitations have been achieved via \emph{spin-orbit resonances}, where
the spin and orbital precession frequencies of the planet become and remain
commensurate as the system evolves (sometimes called \emph{resonance
advection}). Spin-orbit resonances have been studied extensively in solar system
dynamics as \emph{Cassini States} and are responsible for the significant
nonzero obliquities of both Jupiter ($3.1^\circ$) and Saturn ($26.7^\circ$)
\citep{colombo1966, henrard1987, ward2004I, ward_jupiter}. Recent work has
attempted to apply this mechanism to the ice giants as well
\citep{hamilton_tilting_ice}.

Producing exoplanet obliquities via spin-orbit resonances was the subject of
\citet{millholland_disk}, where a dissipating protoplanetary disk causes
resonance capture and advection. In their work, they begin with a planet whose
spin is nearly aligned with the angular momentum vector of the protoplanetary
disk. After the disk dissipates, the resulting planetary obliquity is generally
substantial, reaching $90^\circ$ for what they characterize as adiabatic
resonance crossings.

However, their study was limited to small initial misalignment angles between
the spin vector of the planet and the orbital angular momentum of the
protoplanetary disk. Such an assumption is intuitive, since the planet forms
from the disk, but prospects for it to be violated exist. In particular, giant
and planetesimal impacts are expected to be abundant during planet formation and
may perturb the spin axis of the protoplanet from its primordial value
\citep{yalinewich2019atmospheric, schlichting2015atmospheric}. Furthermore, the
abundance of compact multi-planet systems provides support for possible
scattering between close-in, newly-formed planets \citep{usp_compact1,
usp_review}. It is conceivable that such processes could generate significant
spin-disk misalignment within the lifetime of the disk.

In our work, we provide an analytical framework for understanding the final
planetary obliquity for arbitrary initial spin-disk misalignment angles. Our
analytical work and independent numerical simulations are in agreement with
\citet{millholland_disk} but provide a more complete characterization of
possible behavior.

The paper is organized as follows. In \autoref{s:eq}, we review relevant
spin-orbit dynamics and present the equations studied in the rest of the paper.
In \autoref{s:ad} and \autoref{s:nonad}, we study the evolution of the system
when the disk either dissipates adiabatically or non-adiabatically respectively.
Exhaustive yet simple analytical arguments are able to explain completely the
numerical results in both limits. We discuss potential consequences in
\autoref{s:disc}. Our primary physical results consist of
\autoref{fig:ad_ensemble} in the adiabatic limit and
\autoref{fig:3_ensemble_05_15} in the non-adiabatic limit. Longer mathematical
calculations are relegated to the appendicies, including a leading-order
estimate of final planetary obliquities given small initial spin-disk
misalignment angles in \autoref{s:ad_approx}.

@@

Much of this treatment parallels that of \citet{anderson2018teeter} and
\citet{millholland_disk}. We consider an oblate star of mass $M_\star$ hosting a
planet of mass $M_p$, radius $R_p$, at semimajor axis $a_p$, and with spin
angular frequency $\Omega_p$, and a protoplanetary disk of mass $M_d$ at some
characteristic distance $r_d$ from the host star \citep[see][for a power-law
disk profile]{millholland_disk}. Denote $\hat{s} \equiv \vec{S} / S$ spin unit
vector of planet, $\hat{l} \equiv \vec{L} / L$ orbital angular momentum unit
vector of planet, and $\hat{l}_d \equiv \vec{L}_d / L_d$ angular momentum of the
surrounding disc. We approximate $S \ll L \ll L_d$, so $\hat{l}_d$ is
approximately constant and $\hat{l}$ experiences negligible backreaction torque
from $\hat{s}$. The quadrupole-order secular equations of motion for $\hat{s}$,
$\hat{l}$ and $\hat{l}_d$ then become very simple
\begin{align}
    \rd{\hat{s}}{t} &= \omega_{sl} \p{\hat{s} \cdot \hat{l}}
        \p{\hat{s} \times \hat{l}},\label{eq:dsdt}\\
    \rd{\hat{l}}{t} &= \omega_{ld}\p{\hat{l} \cdot \hat{l}_d}
        \p{\hat{l} \times \hat{l}_d}\label{eq:dldt}.
\end{align}
where
\begin{align}
    \omega_{sl} &\equiv \frac{3k_{qp}}{2k_p} \frac{M_\star}{m_p}
        \p{\frac{R_p}{a_p}}^3 \Omega_p,\label{eq:wsl}\\
    \omega_{ld} &\equiv \frac{3M_d}{4M_\star}\p{\frac{a_p}{r_d}}^3 n
        .\label{eq:wld}
\end{align}
% = 1/(152600 yr) (kp = 6kqp, Mstar = msun, mp = mJ, Rp = RJ, ap = AU, Wp = 2pi
% / (10 days))
% = 1 / (573362 yr) (Md/Mstar = 0.01, ap/rd = 1/30)
In \autoref{eq:wsl}, $k_{qp}$ and $k_p$ are related to the planet's quadrupole
moment and moment of inertia respectively \citep[see][]{lai2018}, and $n \equiv
\sqrt{GM_\star/a_p^3}$ is the planet's orbital mean motion. The three angles
formed by the three vectors are denoted
\begin{align}
    \theta &\equiv \arccos \hat{s} \cdot \hat{l},\\
    \theta_{sd} &\equiv \arccos \hat{s} \cdot \hat{l}_d,\\
    I &\equiv \arccos \hat{l} \cdot \hat{l}_d.
\end{align}

Let us transform to frame rotating about $\hat{l}_p$ with frequency $g \equiv
-\omega_{ld}\cos I < 0$, then $\hat{l}$ is fixed. In this frame, $\hat{l}_p$
also remains fixed and $\hat{s}$ evolves as:
\begin{equation}
    \rd{\hat{s}}{t} = \alpha \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        + g\p{\hat{s} \times \hat{l}_d},
\end{equation}
where $\alpha \equiv \omega_{sl} > 0$ follows standard notation. By defining
$\eta$ and nondimensionalized time $\tau$ as
\begin{align}
    \eta \equiv{}& \frac{\abs{g}}{\alpha}\label{eq:eta},\\
        ={}& 3.75 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p}{0.001 M_\odot}}
            \p{\frac{R_p}{R_J}}^{-3}
            \p{\frac{\Omega_p}{2\pi/\p{10\;\mathrm{days}}}}^{-1}\nonumber\\
        &\p{\frac{M_d}{0.01 M_{\odot}}}
            \p{\frac{a_p}{1\;\mathrm{AU}}}^{9/2}
            \p{\frac{r_d}{30 \;\mathrm{AU}}}^{-3}
            \p{\frac{M_\star}{1M_{\odot}}}^{-3/2}
            \cos I,\\
    \tau \equiv{}& \alpha t,
\end{align}
we can non-dimensionalize the equation of motion as
\begin{equation}
    \rd{\hat{s}}{\tau} = \p{\hat{s} \cdot \hat{l}}
            \p{\hat{s} \times \hat{l}}
        - \eta\p{\hat{s} \times \hat{l}_d}. \label{eq:dsdt_base}
\end{equation}

Following~\cite{millholland_disk}, we allow $g$ to vary in time owing to a
disk with decaying mass
\begin{equation}
    M_d(t) = M_d(0)e^{-t/t_d}.
\end{equation}
As $\alpha$ is held constant, this equates to $\eta$ decaying as $\rd{\eta}{t} =
-\frac{\eta}{t_d}$ as well. We nondimensionalize:
\begin{align}
    \epsilon \equiv{}& \frac{1}{\alpha t_d},\\
        ={}& 0.015 \p{\frac{k_p}{6k_{qp}}}
            \p{\frac{m_p/M_\star}{0.001}}
            \p{\frac{a_p}{1\;\mathrm{AU}}}^3\nonumber\\
        &\p{\frac{R_p}{R_J}}^{-3}
            \p{\frac{\Omega_p}{2\pi/\p{10\;\mathrm{days}}}}
            \p{\frac{t_d}{10\;\mathrm{Myr}}}^{-1},\\
    \rd{\eta}{\tau} ={}& -\epsilon \eta.\label{eq:deta_dt}
\end{align}
\autoref{eq:dsdt_base} and \autoref{eq:deta_dt} together constitute our system
of study.

@@

Spin states satisfying $\rd{\hat{s}}{t} = 0$ are referred to as \emph{Cassini
States} (CSs). There are either two or four solutions, depending on the values
of $\eta, I$. Following standard convention and nomenclature (see
\autoref{fig:cs_locs}), CSs 1, 3, 4 have $\phi = 0, \theta < 0$ and CS2 has
$\phi = \pi, \theta > 0$. Of these, CSs 1, 2, 3 are stable while CS4 is unstable
(see \autoref{s:local_dynamics}). When $\eta < \eta_c$, where
\begin{equation}
    \eta_c \equiv \p{\sin^{2/3}I + \cos^{2/3}I}^{-3/2},
\end{equation}
all four CSs exist, and when $\eta > \eta_c$, only CSs 2, 3 exist
\citep{henrard1987,ward2004I}. The CS obliquities as a function of $\eta$ are
shown in \autoref{fig:cs_locs}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/2_cs_locs.png}
    \caption{Cassini state obliquities as a function of $\eta$. Note that
    $\theta \in [-\pi, \pi]$ is the traditional definition of the polar angle
    \citep[see e.g.][]{colombo1966,peale1969,henrard1987}.}\label{fig:cs_locs}
\end{figure}

@@

When $\eta < \eta_c$, CS4 exists and is a saddle point. The two trajectories
originating and ending at CS4 are the only two infinite-period orbits in phase
space. Together, these two critical trajectories are referred to as the
\emph{separatrix} and divide phase space into three zones. The separatrix, three
zones, and their relations to the CSs are shown via a contour plot of the
Hamiltonian of the system
\begin{equation}
    \mathcal{H} = -\frac{1}{2}\p{\hat{s} \cdot \hat{l}}^2
            + \eta \p{\hat{s} \cdot \hat{l}_d}.\label{eq:H}
\end{equation}
Such a contour plot is given in \autoref{fig:eq_1contours}. Note that
trajectories in zone $II$ librate about CS2 while those in zones $I$/$III$
circulate.
\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{../initial/0_eta/1contours_flip.png}
    \caption{Contour plot of $\mathcal{H}\p{\phi, \cos \theta}$ as given in
    \autoref{eq:H}, where warmer colors denote more positive values. The black
    solid line is the separatrix, which only exists for $\eta < \eta_c$. The
    three zones, divided by the separatrix, are labeled. The Cassini states are
    labeled and have the same colors as \autoref{fig:cs_locs}. The interior of
    the separatrix, shaded in grey, is formally only defined for $\eta <
    \eta_c$, but we may identify the phase space trajectories that flow into
    zone $II$ when evolved forward in time; this is the shaded region in the top
    left plot, bounded by the dotted black line.}\label{fig:eq_1contours}
\end{figure*}

The unsigned areas of the three zones are known exactly
\citep{henrard1987,ward2004I}. Defining
\begin{align*}
    z_0 &= \eta\cos I, &
    \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
    \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
        \chi^2 \cos^2\theta_4 + 1},&
    T &= 2\chi \frac{\cos \theta_4}{
        \chi^2 \cos^2\theta_4 - 1},\\
\end{align*}
the areas for $\eta < \eta_c$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_I &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in \autoref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_c$, but a natural extension
exists: evolve an initial point $p$ under adiabatic decrease of $\eta > \eta_c$
until the separatrix appears at $\eta = \eta_c$, then identify $p$ with the zone
it is in at $\eta_c$. Since phase space area is conserved under adiabatic
evolution, this implies $A_i\p{\eta > \eta_c} = A_i(\eta_c)$. This extension is
also reflected in the dashed boundary in \autoref{fig:eq_1contours}.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{Plot of fractional areas of each of the zones $A_{i}(\eta) / 4\pi$
    as given by \autoref{se:area_ward}. Dotted lines correspond to small $\eta$
    approximations discussed in \autoref{s:ad_approx}. Dashed lines for $\eta >
    \eta_c$ effective values of $A_{II}, A_{III}$ for $\eta > \eta_c$, denoting
    the area of points that would flow into either area under adiabatic decrease
    of $\eta$ from $\eta > \eta_c$.}\label{fig:eq_areas}
\end{figure}

@@

One criterion for adiabaticity is that the separatrix crossing timescale be
longer than the the resonant libration period \citep{ward2004II}. The resonant
libration period for a nearly separatrix-crossing orbit can be approximated by
the libration period near CS2, resulting in constraint
\begin{align}
    -\at{\frac{1}{\eta}\rd{\eta}{\tau}}_{cross} = \epsilon &\lesssim
            \frac{1}{T_{lib}},\\
        &\lesssim \frac{1}{2\pi}\sqrt{\eta_\star\sin I \sin \theta_2
            \p{1 + \eta_\star \sin I \csc^3 \theta_2}},
            \label{eq:ad_constr}
\end{align}
where $\theta_2$ is the obliquity at CS2. This formula differs from that
presented in \citet{millholland_disk}, where the $\csc^3\theta_2$ term is
neglected and a square root is missing. See \autoref{s:local_dynamics} for a
detailed derivation of $T_{lib}$ and comparison to existing work.

@@

We evenly sample $101$ values of $\theta_{sd, i}$, and for each $\theta_{sd, i}$
value, we pick $101$ $\hat{s}$ initial conditions approximately on the ring
fixed by $\theta_{sd, i}$\footnote{For each $\theta_{sd, i}$ we choose an
initial condition on the libration cycle going through $\p{\theta_2 +
\theta_{sd, i}, \phi_2}$ (where $\p{\theta_2, \phi_2}$ are the coordinates of
CS2), such that all initial conditions for a given $\theta_{sd, i}$ enclose the
same initial phase space; see \autoref{sss:a_evo}.}. To be concrete, we choose
$\eta_i = 10\eta_c$. For each initial condition, we evolve
Equations~\ref{eq:dsdt_base} and~\ref{eq:deta_dt} until $\eta = 10^{-5}$. For
such small $\eta$, $\hat{s}$ precesses about $\hat{l}$ with uniform final
obliquity $\theta_f$. This transfer function $\theta_f\p{\theta_{sd, i}}$ is our
primary result, and is shown for $I = 5^\circ, 10^\circ, 20^\circ$ in
\autoref{fig:ad_ensemble}, \autoref{fig:3_ensemble_10_35}, and
\autoref{fig:3_ensemble_20_35} respectively. The blue dots represent the results
of the simulation, which are in excellent agreement with the colored tracks.
These colored tracks are calculated semi-analytically using methods explained in
the following section.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_05_35.png}
    \caption{Top: $\theta_{f}\p{\theta_{sd, i}}$, overlaid with semi-analytic
    predictions of the $\theta_{ f}$ for each of the four nontrivial
    dynamical tracks in colored lines. Black dotted lines correspond to
    small-$\theta_{sd, i}$ predictions derived in \autoref{s:ad_approx}.
    Bottom: Semi-analytic probabilities of each of the dynamical tracks for each
    $\theta_{sd, i}$. The five regimes of $\theta_{sd, i}$ values demarcated by
    the vertical dashed black lines correspond to the five regimes of $A_i$
    calculated in \autoref{ss:zone_transitions}. Black dotted lines correspond to
    simple analytic estimates given in
    \autoref{s:ad_approx}.}\label{fig:ad_ensemble}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_10_35.png}
    \caption{Same as \autoref{fig:ad_ensemble} but for $I =
    10^\circ$ and with fewer annotations.}\label{fig:3_ensemble_10_35}
\end{figure}
\begin{figure}
    \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3_ensemble_20_35.png}
    \caption{Same as \autoref{fig:3_ensemble_10_35} but for $I = 20^\circ$. The
    $III \to II \to I$ track is very short since the maximum $A_{II}$ is very
    near $\eta_c$ for $I = 20^\circ$.}\label{fig:3_ensemble_20_35}
\end{figure}
